%
% Acknowledgements
\chapter{Acknowledgements}%

You never write a master thesis alone. There are a great deal of people to thank for getting me to complete this work, not only in an academic sense but also from the moral support side.\par 

First of all I would like to thank my supervisor Manon Kok for taking the time to tackle this master thesis with me full on. You have helped me understand topics I thought would be very difficult to wrap my head around. You have given good advice, but have also given me the freedom to discover things on my own, allowing me to learn in the process. \par 

Then there is the sensor fusion group of master students with whom problems have been shared, potential solutions have been discussed and thesis sections have been scrutinized. Thanks to all of you: Frida, Thomas, Evan, Daan, and Stefan!

The support group around has also been essential. This range from friends I asked to give feedback on thesis sections to family asking what I have been doing for the last year, and always being impressed, even though they do not really understand it.

I especially want to thank Julia Brevoord, who has been patient with me the whole time and has brought moments of stress into perspective. For also enduring countless requests for feedback. At this point I think you may understand my thesis subject better than I do.

Last but not least, my eternal thanks to my parents, Hein and Kiek van Ingen, without whose support non of this would have been possible.



\vspace*{15mm}

Delft, University of Technology \hfill \mscname \\
\mscdate